Uncovering underlying high-level musical content in the temporal domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music embeddings with applications to MIR downstream tasks

### 2023-05-15 evaluation/subsections/beyond dsp.tex
Updates from Overleaf

Let's pick a core MIR task as a sample to display such complexity: Music Structure Analysis (MSA), an interdisciplinary field that aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. However, due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

The main principles of MSA were initially defined as homogeneity, novelty, and repetition, with the addition of regularity. The checkerboard kernel technique is a simple and effective method for MSA based on the homogeneity principle. The kernel, with a checkerboard-like structure, is convolved over the main diagonal of a Self-Similarity Matrix (SSM) such as $S_{ij}$ is the similarity between time frames $i$ and $j$, $\vec{x}_i$ is the vector representation of time frame $i$, and $\left| \vec{x}_i \right|$ is the Euclidean norm of vector $\vec{x}_i$. The dot product of two vectors $\vec{x}_i$ and $\vec{x}_j$ is denoted by $\vec{x}_i \cdot \vec{x}_j$.

\input{introduction/subsections/equations/SSM equation}

This yields a novelty curve highlighting sudden changes in the selected musical features from which to extract the segment boundaries.

\input{introduction/subsections/equations/Novelty equation}

Mathematical operations are used on predefined principles such as homogeneity and repetition to analyze a musical structure. It explores features extracted from a musical signal and identifies peaks in a novelty curve to extract segment boundaries. In contrast, human aural skills rely on the perception and interpretation of music by human listeners based on cultural and academic background. They can capture subtle nuances of musical structure, but those still are subjective and can vary across listeners.

Combining technical tasks, aural skills, and music perception is challenging due to the complexity and subjectivity of musical perception, discrepancies between the features extracted by DSP techniques and the features perceived by humans, and the variability of musical signals and individual differences in musical perception.

Purely mathematical approaches may not always reflect how humans perceive music and may not capture the subtle variations in timing or harmonic relationships critical to musical perception. Variability in musical signals due to instrumentation, genre, and performance style further complicates the development of these techniques that can be generalized to different musical contexts.

@@

Features extracted from audio files play a vital role in characterizing and differentiating sound and music. One such feature, known as the \textit{Embeddiogram}, offers a unique and powerful representation of an audio signal. The Embeddiogram is derived by applying a pre-trained neural network model to windowed segments of an audio signal, generating a sequence of embedding vectors. These embedding vectors collectively form a high-dimensional description of the audio signal's structure and content.

@@

>>>>

While the usefulness of deep audio embeddings can be evaluated in several tasks, we have chosen a core MIR task for its popularity and complexity: Music Structure Analysis (MSA). This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. However, due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

The main principles of MSA were initially defined as homogeneity, novelty, and repetition, with the addition of regularity.

@@

Features extracted from audio files play a vital role in characterizing and differentiating sound and music. One such feature, the \textit{Embeddiogram}, offers a unique and powerful representation of an audio signal. The Embeddiogram is derived by applying a pre-trained neural network model to windowed segments of an audio signal, generating a sequence of embedding vectors. These embedding vectors collectively form a high-dimensional description of the audio signal's structure and content.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/images/embeddiogram_SALAMI_track_2.png}
    \caption[Embeddiogram]{Caption}
    \label{fig:embeddiogram}
\end{figure}

### 2023-05-15 introduction/Subsections/Assumptions.tex
Updates from Overleaf

This thesis posits the existence of high-level musical concepts invariant to specific attributes such as tonal quality, tempo, key, orchestration, and signal noise that evolve and unfolds through time. This idea parallels the nature of the symbolic domain (sheet music), which maintains its essence despite being interpreted through diverse performances using various instruments and styles. While a single waveform might have a limited number of sheet music representations that adequately capture its musical components, a single piece can be interpreted in infinite ways through different instruments, voices, tempos, and interpretations. The distinct tonal quality of a waveform, heavily influenced by the specific timbre of the instrument or production technology, adds complexity to the transcription of its sonic properties into accurate sheet music notation.

>>>>

This thesis posits the existence of high-level musical concepts invariant to specific attributes such as sound quality, tempo, key, orchestration, and signal noise that evolve and unfolds through time. This idea parallels the nature of the symbolic domain (sheet music), which maintains its essence despite being interpreted through diverse performances using various instruments and styles. While a single waveform might have a limited number of sheet music representations that adequately capture its musical components, a single piece can be interpreted in infinite ways through different instruments, voices, tempos, and interpretations. The distinct tonal quality of a waveform, heavily influenced by the specific timbre of the instrument or production technology, adds complexity to the transcription of its sonic properties into accurate sheet music notation.

### 2023-05-15 introduction/Subsections/Motivation.tex
Updates from Overleaf

I advocate for exposure as the primary method of learning music, emphasizing the importance of immersing oneself in various musical experiences to foster a holistic understanding of music as an art form. Individuals can cultivate well-rounded musical expertise by engaging with different styles, genres, techniques, and learning approaches, ultimately promoting creativity and experimentation.

@@

@@

\begin{figure}[ht]
\includegraphics[clip,width=\columnwidth]{figures/schenkerian analysis/SchubertOp4no3.png}% 
\caption[Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert.]{\small{Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert. Passage's original score, the schenkerian unfolding of the melody, the chord degrees analysis, and their tonal function.}}
\label{fig:Wandrers Nachtlied, Op. 4, D. 224}
\end{figure}

>>>>

@@

Purely mathematical approaches may not always reflect how humans perceive music and may not capture the subtle variations in timing or harmonic relationships critical to musical perception.

@@

\begin{figure}[ht]
\includegraphics[clip,width=\columnwidth]{figures/schenkerian analysis/SchubertOp4no3.png}% 
\caption[Excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert.]{\small{Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert. Passage's original score, the schenkerian unfolding of the melody, the chord degrees analysis, and their tonal function.}}
\label{fig:Wandrers Nachtlied, Op. 4, D. 224}
\end{figure}

### 2023-05-15 introduction/Subsections/Previous work.tex
Updates from Overleaf

\item The utilization of multi-level music segmentation through deep audio embeddings has replaced manually engineered features and demonstrated significant performance improvement over existing approaches \cite{SalamonDeepSegmentation}.

\item The enhancement of environmental sound classification, suggesting the potential for further improvements through class-specific augmentations \cite{Kim2020OneStrategies}.

@@

Returning to musicological influences, the concept of the middle ground can be compared to the mesostructure in music. The middle ground, analogous to the mesostructure, represents the intermediate-level musical patterns that bridge the micro and macro structures. This element is often overlooked when applying deep learning to music. As emphasized by \cite{Mesostructures2023}, models that address this mesostructure level could significantly enrich tasks such as music analysis, composition, and retrieval, providing a more holistic understanding of music similar to the insights gained from analyzing the middle-ground in Schenkerian analysis. \cite{Introduction_to_Schenkerian_Analysis}

>>>>

\item The utilization of multi-level music segmentation through deep audio embeddings to replace manually engineered features \cite{SalamonDeepSegmentation}.

\item The enhancement of environmental sound classification \cite{Kim2020OneStrategies}.

@@

On top of that and returning to musicological influences, the middle ground in Schenkerian analysis can be compared to the mesostructure in music. Analogous to the mesostructure, it represents the intermediate-level musical patterns that bridge the micro and macro structures. This element is often overlooked when applying deep learning to music. As emphasized by \cite{Mesostructures2023}, models that address this mesostructure level could significantly enrich tasks such as music analysis, composition, and retrieval, providing a more holistic understanding of music similar to the insights gained from analyzing the middle-ground in Schenkerian analysis. \cite{Introduction_to_Schenkerian_Analysis}

### 2023-05-15 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

@@

\subsubsection{Self-Supervised Learning (SSL)}

@@

\subsubsection{Contrastive Learning of Musical Representations (CLMR)}

@@

Triplet Siamese Networks, an extension of the Siamese Network architecture, involve comparing three input instances instead of two, aiming to learn an embedding space where similar instances are close and dissimilar instances are distant. The input consists of an anchor, a positive instance (same class as the anchor), and a negative instance (different class). Each instance is processed by an identical subnetwork with shared architecture, weights, and hyperparameters.

@@

\subsection{XXX}
\subsection{XXX}

\subsection{Encoder}
\subsubsection{SampleCNN}

@@

\item ReLU activation. \input{figures/equations/ReLU}

@@

\begin{tabular}{|c|c|c|}
\hline
\textbf{Layer} & \textbf{Input Channels} & \textbf{Output Channels} \\
\hline
1 & 128 & 128 \\
\hline
2 & 128 & 128 \\
\hline
3 & 128 & 256 \\
\hline
4 & 256 & 256 \\
\hline
5 & 256 & 256 \\
\hline
6 & 256 & 256 \\
\hline
7 & 256 & 256 \\
\hline
8 & 256 & 256 \\
\hline
9 & 256 & 512 \\
\hline
\end{tabular}

@@

The output of the final convolutional layer is passed through an average pooling operation across the temporal dimension (dim=2).

@@

We have adapted the original architecture to our purpose, which consists of the final average pooling step.

\begin{equation}
y_i = \frac{1}{N} \sum_{j=1}^{N} x_{ij}
\end{equation}

$y_i$ represents the output tensor after the average pooling operation, $x_{ij}$ represents the elements in the 2D input tensor, $N$ is the number of elements along the second dimension of the input tensor, $i$ iterates over the first dimension of the tensors, and $j$ iterates over the second dimension of the input tensor.

>>>>

I advocate for exposure as the primary method of learning music, emphasizing the importance of immersing oneself in various musical experiences to foster a holistic understanding of music as an art form. Individuals can cultivate well-rounded musical expertise by engaging with different styles, genres, techniques, and learning approaches, ultimately promoting creativity and experimentation.

@@

\subsection{Self-Supervised Learning (SSL)}

@@

\subsection{Contrastive Learning of Musical Representations (CLMR)}

@@

Triplet Siamese Networks, an extension of the Siamese Network architecture, involve comparing three input instances instead of two, aiming to learn an embedding space where similar instances are close and dissimilar instances are distant. The input consists of an anchor, a positive instance (same class as the anchor), and a negative instance (different class). An identical subnetwork with shared architecture, weights, and hyperparameters processes each instance.

@@

\subsection{Convolutional Neural Network (CNN)}

A Convolutional Neural Network (CNN) is a deep-learning neural network. The architecture of a CNN is composed of several layers, including:

\begin{itemize}

\item \textbf{Convolutional layers:} These layers apply a convolution operation to the input data. The convolutional layers are designed to effectively learn local patterns and features in the audio data \footnote{For a visual explanation and further understanding of the convolution operation, please refer to this URL: \url{https://youtu.be/KuXjwB4LzSA}}.

Let's consider a 1D input sequence or signal represented as a vector $I = [i_{i}]{n}$ and a filter $K = [k_{j}]_{p}$, where $n$ is the length of the input sequence and $p$ is the length of the filter. The convolution operation is defined as:

\begin{equation}
(I * K){k} = \sum{j=1}^{p} i_{k+j-1}k_{j}
\end{equation}

Here, $k$ ranges from $1$ to $n-p+1$. The resulting feature vector after applying the convolution is denoted by $(I * K)_{k}$.

The convolution operation slides the kernel $K$ over the input sequence $I$, performing an element-wise multiplication and sum of the local region centered at element $k$ to compute the value of each element in the output feature vector, $(I * K)_{k}$. This process is repeated for each component $k$, resulting in a new feature vector that captures local patterns in the input sequence.

In simple terms, the filter or kernel is slid across the 1D input data. For each position, we perform an element-wise multiplication of the filter with the portion of the input it currently covers and sums the results. This value becomes the corresponding entry in the output. The process is repeated for each position the filter can occupy.

\item \textbf{Pooling layers:} These layers downsample the data, reducing the input dimensions while retaining important information. A joint pooling operation is the max-pooling, which can be defined as follows: given an input matrix $I$ and a pooling window of size $m \times n$, the max-pooling operation selects the maximum value within each non-overlapping $m \times n$ region of the input matrix. This operation effectively reduces the input's dimensions while retaining the essential information in each area.
\vspace*{3mm}

\item \textbf{Fully connected layers:} These layers connect every neuron in one layer to every neuron in another, allowing the network to learn non-linear combinations of the features discovered in the previous layers. Mathematically, a fully connected layer can be represented as $Y = XW + b$, where $X$ is the input matrix, $W$ is the weight matrix, $b$ is the bias vector, and $Y$ is the output matrix.
\vspace*{3mm}

\item \textbf{Output layer:} The output layer produces the final predictions of the network.

\end{itemize}

Compared to a traditional neural network, CNNs are more computationally efficient and have less number of parameters to train. This makes them more feasible for large-scale datasets and real-world problems. On top of that, one of the main advantages of using CNNs for audio analysis is that they can automatically and adaptively learn temporal hierarchies of features from audio data, which traditional audio processing methods may not do effectively.

%Colour CNN
\input{figures/neural networks/colour CNN}

\subsection{Encoder: SampleCNN}

@@

\item ReLU activation. The Rectified Linear Unit (ReLU) activation function is a simple yet highly effective non-linear function used in neural networks. It helps introduce non-linearity to the model and accelerates training due to its computational efficiency.

@@

@@

The original model has been modified slightly to serve our specific needs: the output from the final convolutional layer is now subjected to an average pooling operation across the temporal dimension.

This adaptation of the original architecture incorporates an average pooling step at the end.

\begin{figure}[ht]
\centering
\begin{equation}
y_i = \frac{1}{N} \sum_{j=1}^{N} x_{ij}
\end{equation}
\caption{\small{Representation of the Average Pooling operation. Here, $y_i$ is the output tensor after the average pooling process. Each element $x_{ij}$ is part of the 2D input tensor. $N$ is the count of elements in the second dimension of the input tensor, while $i$ and $j$ are used to iterate over the first and second dimensions of the tensors, respectively.}}
\end{figure}

@@
